---
title: "5010 Project"
author: "Qichen He, Tongyuan Lu, Chenxi Wang, Yufeng Xue"
output: html_document
---

Suppose we have $100,000 at hand, and we would like to find the optimal stock portfolio to invest in on January 1, 2020. The minimum variance portfolio (MVP) minimizes risks, and the tangency portfolio (TP) is the one with the highest Sharpe's ratio. Our goal is to identify the superior portfolio based on the following assumptions: the risk-free rate is 1.59%, and stock shares can have decimal places.

The portfolio consists of five stocks from the Dow Jones Industrail Average:  Apple Inc.(AAPL), The Boeing Company (BA), Gilead Sciences, Inc.(GILD), The Goldman Sachs Group, Inc.(GS) and 3M Company (MMM). In order to reduce investment risks, these stocks are selected from five different sectors. It also contains five bonds: Metropolitan West High Yield Bond Fund(MWHYX), iShares Short-Term Corporate Bond (IGSB), iShares Intermediate-Term Corporate Bond (IGIB), Vanguard Short-Term Corporate Bond (VCSH) and Vanguard Intermediate-Term Corporate Bond (VCIT). We use yahoo finance data from January 02, 2018 to December 31, 2019 for our analysis. 


```{r, message = FALSE}
library(readr)
library(quadprog)
library(dplyr)
library(ggplot2)
```


```{r, message = FALSE}
# load stock data
AAPL <- read.csv("~/Desktop/5010/5010 Project/AAPL (1).csv")
BA <- read.csv("~/Desktop/5010/5010 Project/BA.csv")
GS <- read.csv("~/Desktop/5010/5010 Project/gs.csv")
MMM <- read.csv("~/Desktop/5010/5010 Project/mmm.csv")
GILD <- read.csv("~/Desktop/5010/5010 Project/gild.csv")

# load bond data
MWHYX <- read.csv("~/Desktop/5010/5010 Project/MWHYX.csv")        
IGSB <- read.csv("~/Desktop/5010/5010 Project/IGSB.csv")          
IGIB <- read.csv("~/Desktop/5010/5010 Project/IGIB.csv") 
VCSH <- read.csv("~/Desktop/5010/5010 Project/VCSH.csv") 
VCIT <- read.csv("~/Desktop/5010/5010 Project/VCIT.csv")

adjClose <- data.frame(AAPL_AC = AAPL$Adj.Close, 
                       BA_AC = BA$Adj.Close, 
                       GS_AC = GS$Adj.Close, 
                       MMM_AC = MMM$Adj.Close, 
                       GILD_AC = GILD$Adj.Close,
                       MWHYX_AC = MWHYX$Adj.Close,
                       IGSB_AC = IGSB$Adj.Close,
                       VCSH_AC = VCSH$Adj.Close,
                       IGIB_AC = IGIB$Adj.Close,
                       VCIT_AC = VCIT$Adj.Close)
```

# MVP and TP (short-selling allowed)

```{r}
# MVP, short-selling allowed
ndays <- nrow(adjClose)
nstocks <- ncol(adjClose)
return <- adjClose[2:ndays,]/adjClose[1:(ndays-1),] - 1
mu <- colMeans(return)
sigma <- cov(return)

m <- 200 # no. of points to evaluate
muP <- seq(.0001,.0015, length=m) # target portfolio return
sdP <- rep(0, length(muP)) # sd of portfolio return
weight <- matrix(0, nrow=m, ncol=nstocks) # storage for portfolio weights
for (i in 1:length(muP)) { # find the optimal portfolios
  result <- solve.QP(Dmat = 2*sigma, 
                     dvec = rep(0, nstocks), 
                     Amat = cbind(rep(1,nstocks), mu), 
                     bvec = c(1, muP[i]), 
                     meq = 2)
  sdP[i] <- sqrt(result$value)
  weight[i,] <- result$solution
}

GMP <- which.min(sdP) # global minimum point
# efficient frontier
plot(sdP[GMP:m], muP[GMP:m], type="l", 
     ylim = c(min(muP)-0.001, max(muP)+0.001), xlim = c(0, max(sdP)+0.001), 
     lwd = 3, col = "red", xlab = "SD of portfolio return",
     ylab = "mean of portfolio return")
points(sdP[1:(GMP-1)], muP[1:(GMP-1)], type="l",
       lty = 2, lwd = 3, col = "red")
legend("bottomright", legend=c("Efficient Frontier (With Short Sale)"),
       col=c("red"), lty=1)

mvr_weights <- weight[GMP, ]
mvr_weights # weights for AAPL, BA, GS, MMM, GILD, MWHYX, IGSB, VCSH, IGIB, VC respectively
mvr_return <- muP[GMP]
mvr_return
mvr_sd <- sdP[GMP]
mvr_sd
```



```{r}
# Tangency portfolio, short-selling allowed
muf = 1.59/36500  # Add risk-free asset, using r = 1.59%
sharpe_ratio = max((muP - muf)/sdP) # Find the largest slope
idx = which.max((muP - muf)/sdP)
plot(sdP[GMP:m], muP[GMP:m], type="l", 
     ylim = c(min(muP)-0.001, max(muP)+0.001), xlim = c(0, max(sdP)+0.001), 
     lwd = 3, col = "red", xlab = "SD of portfolio return",
     ylab = "mean of portfolio return")
points(sdP[1:(GMP-1)], muP[1:(GMP-1)], type="l",
       lty = 2,lwd = 3, col = "red")
#points(sqrt(diag(sigma)), mu, pch = 4) # label
#text(sqrt(diag(sigma))+0.002, mu, c("CAT","IBM","MSFT")) # text label
abline(v = 0)
points(0, muf, pch = 4) # Plot risk-free asset
text(0.00025, muf-0.00025, c("F"))
points(sdP[idx], muP[idx], pch = 4) # Plot tangency portfolio
text(sdP[idx], muP[idx]+0.00025, c("T"))
lines(x = c(0, sdP[idx]), y = c(muf, muP[idx]), lwd = 3, col = "blue")
legend("bottomright", legend=c("Efficient Frontier (With Short Sale)"),
       col=c("red"), lty=1)

t_weights <- weight[idx, ]
t_weights # weights for AAPL, BA, GS, MMM, GILD, MWHYX, IGSB, VCSH, IGIB, VC respectively
t_return <- muP[idx]
t_return
t_sd <- sdP[idx]
t_sd
```


# MVP and TP short-selling NOT allowed

```{r}
# MVP, short sale NOT allowed
muP_noSS <- seq(min(mu), max(mu),length=m) # target portfolio return
sdP_noSS <- rep(0, length(muP_noSS)) # sd of portfolio return
weight_noSS <- matrix(0, nrow=m, ncol=nstocks) # storage for portfolio weights
for (i in 1:length(muP_noSS)) { # find the optimal portfolios
  result <- solve.QP(Dmat = 2*sigma,
                    dvec = rep(0, nstocks),
                    Amat = cbind(rep(1,nstocks), mu, diag(1,nstocks)),
                    bvec = c(1, muP_noSS[i], rep(0,nstocks)),
                    meq = 2)
  sdP_noSS[i] <- sqrt(result$value)
  weight_noSS[i,] <- result$solution
}


GMP_noSS <- which.min(sdP_noSS) # global minimum point

# efficient frontier when short-selling allowed
plot(sdP[GMP:m], muP[GMP:m], type="l", 
     ylim = c(min(muP)-0.001, max(muP)+0.001), xlim = c(0, max(sdP)+0.001), 
     lwd = 3, col = "red", xlab = "SD of portfolio return",
     ylab = "mean of portfolio return")
points(sdP[1:(GMP-1)], muP[1:(GMP-1)], type="l",
       lty = 2, lwd = 3, col = "red")

# efficient frontier when short-selling NOT allowed
points(sdP_noSS[GMP_noSS:m], muP_noSS[GMP_noSS:m], type="l",
       lwd = 3, col = "blue")
points(sdP_noSS[1:(GMP_noSS-1)], muP_noSS[1:(GMP_noSS-1)], type = "l",
       lty = 2,lwd = 3, col = "blue")
legend("bottomright", legend=c("With Short Sale","No Short Sale"),
       col=c("red", "blue"), lty=1:1)

mvr_noSS_weights <- weight_noSS[GMP_noSS, ]
mvr_noSS_weights # weights for AAPL, BA, GS, MMM, GILD, MWHYX, IGSB, VCSH, IGIB, VC respectively
mvr_noSS_return <- muP_noSS[GMP_noSS]
mvr_noSS_return
mvr_noSS_sd <- sdP_noSS[GMP_noSS]
mvr_noSS_sd
```


```{r}
# Tangency, short-selling NOT allowed
muf = 1.59/36500  
sharpe_ratio_noSS = max((muP_noSS - muf)/sdP_noSS) # Find the largest slope
idx_noSS = which.max((muP_noSS - muf)/sdP_noSS)
plot(sdP_noSS[GMP_noSS:m], muP_noSS[GMP_noSS:m], type="l", 
     ylim = c(min(muP)-0.001, max(muP)+0.001), xlim = c(0, max(sdP)+0.001), 
     lwd = 3, col = "red", xlab = "SD of portfolio return",
     ylab = "mean of portfolio return")
points(sdP_noSS[1:(GMP_noSS-1)], muP_noSS[1:(GMP_noSS-1)], type="l",
       lty = 2,lwd = 3, col = "red")
#points(sqrt(diag(sigma)), mu, pch = 4) # label
#text(sqrt(diag(sigma))+0.002, mu, c("CAT","IBM","MSFT")) # text label
abline(v = 0)
points(0, muf, pch = 4) # Plot risk-free asset
text(0.00025, muf-0.00025, c("F"))
points(sdP_noSS[idx_noSS], muP_noSS[idx_noSS], pch = 4) # Plot tangency portfolio
text(sdP_noSS[idx_noSS], muP_noSS[idx_noSS]+0.00025, c("T"))
lines(x = c(0, sdP_noSS[idx_noSS]), y = c(muf, muP_noSS[idx_noSS]), 
      lwd = 3, col = "blue")
legend("bottomright", legend=c("Efficient Frontier (No Short Sale)"),
       col=c("red"), lty=1)

t_noSS_weights <- weight_noSS[idx_noSS, ]
t_noSS_weights # weights for AAPL, BA, GS, MMM, GILD, MWHYX, IGSB, VCSH, IGIB, VC respectively
t_noSS_return <- muP_noSS[idx_noSS]
t_noSS_return
t_noSS_sd <- sdP_noSS[idx_noSS]
t_noSS_sd
```



```{r}
# calculate returns of each stock, bond and Minimum Variance portfolios, Tangency portfolios
adjClose_return <- adjClose %>%
                   mutate(date = AAPL$Date) %>%
                   select(date, AAPL_AC, BA_AC, GS_AC, MMM_AC, GILD_AC, MWHYX_AC,
                          IGSB_AC, VCSH_AC, IGIB_AC, VCIT_AC) %>%
                   mutate(AAPL_return = (AAPL_AC - lag(AAPL_AC))/lag(AAPL_AC),
                          BA_return = (BA_AC -lag(BA_AC))/lag(BA_AC),
                          GS_return = (GS_AC - lag(GS_AC))/lag(GS_AC),
                          MMM_return = (MMM_AC - lag(MMM_AC))/lag(MMM_AC),
                          GILD_return = (GILD_AC - lag(GILD_AC))/lag(GILD_AC),
                          MWHYX_return = (MWHYX_AC -lag(MWHYX_AC))/lag(MWHYX_AC),
                          IGSB_return = (IGSB_AC -lag(IGSB_AC))/lag(IGSB_AC),
                          VCSH_return =  (VCSH_AC -lag(VCSH_AC))/lag(VCSH_AC),
                          IGIB_return =  (IGIB_AC -lag(IGIB_AC))/lag(IGIB_AC),
                          VCIT_return =  (VCIT_AC -lag(VCIT_AC))/lag(VCIT_AC),
                          mvr_return = mvr_weights[1]*AAPL_return + 
                                       mvr_weights[2]*BA_return +
                                       mvr_weights[3]*GS_return +
                                       mvr_weights[4]*MMM_return +
                                       mvr_weights[5]*GILD_return +
                                       mvr_weights[6]*MWHYX_return +
                                       mvr_weights[7]*IGSB_return +
                                       mvr_weights[8]*VCSH_return +
                                       mvr_weights[9]*IGIB_return +
                                       mvr_weights[10]*VCIT_return,
                          t_return = t_weights[1]*AAPL_return + 
                                     t_weights[2]*BA_return +
                                     t_weights[3]*GS_return +
                                     t_weights[4]*MMM_return +
                                     t_weights[5]*GILD_return +
                                     t_weights[6]*MWHYX_return +
                                     t_weights[7]*IGSB_return +
                                     t_weights[8]*VCSH_return +
                                     t_weights[9]*IGIB_return +
                                     t_weights[10]*VCIT_return)



```


# Calcualte VaR through the historical simulation method

Historical Simulation Method
Calculating VaR (Minimum Variance Portfolio)
Short-Selling Allowed 
```{r}
# VaR: historical simulation (calculate gross return)
var_data_mvr <- adjClose_return %>%
                select(date, AAPL_AC, BA_AC, GS_AC, MMM_AC, GILD_AC, MWHYX_AC,
                          IGSB_AC, VCSH_AC, IGIB_AC, VCIT_AC) %>%
                mutate(AAPL_g_return = AAPL_AC / lag(AAPL_AC),
                       BA_g_return = BA_AC / lag(BA_AC),
                       GS_g_return = GS_AC / lag(GS_AC),
                       MMM_g_return = MMM_AC / lag(MMM_AC),
                       GILD_g_return = GILD_AC / lag(GILD_AC),
                       MWHYX_g_return = MWHYX_AC / lag(MWHYX_AC),
                       IGSB_g_return = IGSB_AC / lag(IGSB_AC),
                       VCSH_g_return = VCSH_AC / lag(VCSH_AC),
                       IGIB_g_return = IGIB_AC / lag(IGIB_AC),
                       VCIT_g_return = VCIT_AC / lag(VCIT_AC)) 

# Assume that we have purchased stocks and bonds based on weights from MVR on Dec 31, 2019, here are the  stock shares
# (to see how good the weights are, we apply the weights on the day that is closest to our intended day of investment)
mvr_shares_dec_31 <- c((100000*mvr_weights[1])/(var_data_mvr$AAPL_AC[503]),
                        (100000*mvr_weights[2])/(var_data_mvr$BA_AC[503]),
                        (100000*mvr_weights[3])/(var_data_mvr$GS_AC[503]),
                        (100000*mvr_weights[4])/(var_data_mvr$MMM_AC[503]),
                        (100000*mvr_weights[5])/(var_data_mvr$GILD_AC[503]),
                        (100000*mvr_weights[6])/(var_data_mvr$MWHYX_AC[503]),
                        (100000*mvr_weights[7])/(var_data_mvr$IGSB_AC[503]),
                        (100000*mvr_weights[8])/(var_data_mvr$VCSH_AC[503]),
                        (100000*mvr_weights[9])/(var_data_mvr$IGIB_AC[503]),
                        (100000*mvr_weights[10])/(var_data_mvr$VCIT_AC[503]))

# calculate portfolio value of dec 31, 2019
dec_31_2019 <- var_data_mvr[503,] %>%
               mutate(portfolio_value = mvr_shares_dec_31[1]*AAPL_AC+
                                        mvr_shares_dec_31[2]*BA_AC +
                                        mvr_shares_dec_31[3]*GS_AC +
                                        mvr_shares_dec_31[4]*MMM_AC +
                                        mvr_shares_dec_31[5]*GILD_AC +
                                        mvr_shares_dec_31[6]*MWHYX_AC +
                                        mvr_shares_dec_31[7]*IGSB_AC +
                                        mvr_shares_dec_31[8]*VCSH_AC +
                                        mvr_shares_dec_31[9]*IGIB_AC +
                                        mvr_shares_dec_31[10]*VCIT_AC) 

## use past gross returns to predict the market value of the portfolio for Jan 1, 2020 (the next day Dec 31, 2019)
  # AAPL_pred: predicted stock price
  # "portfolio_value_pred": prediced portfolio values of Jan 01, 2020 (value of dec 31 * gross returns)
  # "gain_loss_pred": differences between predicted portoflio values and dec 31 portfolio values 
jan_1_pred <- data.frame(AAPL_pred = (var_data_mvr$AAPL_g_return)*dec_31_2019[1,2],
                         BA_pred = (var_data_mvr$BA_g_return)*dec_31_2019[1,3],
                         GS_pred = (var_data_mvr$GS_g_return)*dec_31_2019[1,4],
                         MMM_pred = (var_data_mvr$MMM_g_return)*dec_31_2019[1,5],
                         GILD_pred = (var_data_mvr$GILD_g_return)*dec_31_2019[1,6],
                         MWHYX_pred = (var_data_mvr$MWHYX_g_return)*dec_31_2019[1,7],
                         IGSB_pred = (var_data_mvr$IGSB_g_return)*dec_31_2019[1,8],
                         VCSH_pred = (var_data_mvr$VCSH_g_return)*dec_31_2019[1,9],
                         IGIB_pred = (var_data_mvr$IGIB_g_return)*dec_31_2019[1,10],
                         VCIT_pred = (var_data_mvr$VCIT_g_return)*dec_31_2019[1,11]) %>% 
              mutate(portolio_value_pred = AAPL_pred*mvr_shares_dec_31[1] + 
                                           BA_pred*mvr_shares_dec_31[2] + 
                                           GS_pred*mvr_shares_dec_31[3] + 
                                           MMM_pred*mvr_shares_dec_31[4] + 
                                           GILD_pred*mvr_shares_dec_31[5] +
                                           MWHYX_pred*mvr_shares_dec_31[6] +
                                           IGSB_pred*mvr_shares_dec_31[7] +
                                           VCSH_pred*mvr_shares_dec_31[8] +
                                           IGIB_pred*mvr_shares_dec_31[9] +
                                           VCIT_pred*mvr_shares_dec_31[10],
                     gain_loss_pred = portolio_value_pred - dec_31_2019$portfolio_value) %>%
              na.omit()


# histogram of grain/losses
ggplot(data = jan_1_pred, aes(gain_loss_pred)) + 
  geom_histogram(fill = "deepskyblue3") +
  xlab("Gain/Loss") +
  ylab("Frequency") +
  labs(title = "Histogram of Gain/Loss (Minimum Variance Portfolio)")

# We have a total of 502 predicted gain/losses, and the 99% value at risk for one day is approximately the fifth smallest value

var_mvr_1day <- sort(jan_1_pred$gain_loss_pred)[5]
abs(var_mvr_1day)

# calculate 15 day 99% VaR
var_mvr_5day <- var_mvr_1day*sqrt(5)
abs(var_mvr_5day)
```


Historical Simulation Method
Calculating VaR (Tangency Portfolio)
Short-Selling Allowed 
```{r}
# VaR: historical simulation (calculate gross return)
var_data_TP <- adjClose_return %>%
               select(date, AAPL_AC, BA_AC, GS_AC, MMM_AC, GILD_AC,MWHYX_AC,
                          IGSB_AC, VCSH_AC, IGIB_AC, VCIT_AC) %>%
               mutate(AAPL_g_return = AAPL_AC / lag(AAPL_AC),
                       BA_g_return = BA_AC / lag(BA_AC),
                       GS_g_return = GS_AC / lag(GS_AC),
                       MMM_g_return = MMM_AC / lag(MMM_AC),
                       GILD_g_return = GILD_AC / lag(GILD_AC),
                       MWHYX_g_return = MWHYX_AC / lag(MWHYX_AC),
                       IGSB_g_return = IGSB_AC / lag(IGSB_AC),
                       VCSH_g_return = VCSH_AC / lag(VCSH_AC),
                       IGIB_g_return = IGIB_AC / lag(IGIB_AC),
                       VCIT_g_return = VCIT_AC / lag(VCIT_AC)
                      ) 

# Assume that we have purchased stocks based on weights from TP on Dec 31, 2019, here are the stock shares
  # (notice that some shares are negative)
tp_shares_dec_31 <- c((100000*t_weights[1])/(var_data_TP$AAPL_AC[503]),
                        (100000*t_weights[2])/(var_data_TP$BA_AC[503]),
                        (100000*t_weights[3])/(var_data_TP$GS_AC[503]),
                        (100000*t_weights[4])/(var_data_TP$MMM_AC[503]),
                        (100000*t_weights[5])/(var_data_TP$GILD_AC[503]),
                        (100000*mvr_weights[6])/(var_data_TP$MWHYX_AC[503]),
                        (100000*mvr_weights[7])/(var_data_TP$IGSB_AC[503]),
                        (100000*mvr_weights[8])/(var_data_TP$VCSH_AC[503]),
                        (100000*mvr_weights[9])/(var_data_TP$IGIB_AC[503]),
                        (100000*mvr_weights[10])/(var_data_TP$VCIT_AC[503]))


# calculate portfolio value of dec 31, 2019

dec_31_2019_tp <- var_data_TP[503,] %>%
                  mutate(portfolio_value = 
                                        tp_shares_dec_31[1]*AAPL_AC+
                                        tp_shares_dec_31[2]*BA_AC +
                                        tp_shares_dec_31[3]*GS_AC +
                                        tp_shares_dec_31[4]*MMM_AC +
                                        tp_shares_dec_31[5]*GILD_AC +
                                        tp_shares_dec_31[6]*MWHYX_AC +
                                        tp_shares_dec_31[7]*IGSB_AC +
                                        tp_shares_dec_31[8]*VCSH_AC +
                                        tp_shares_dec_31[9]*IGIB_AC +
                                        tp_shares_dec_31[10]*VCIT_AC) 

## use past gross returns to predict the market value of the portfolio for Jan 1, 2020 (the next day Dec 31, 2019)
  # AAPL_pred: predicted stock price 
  # "portfolio_value_pred": prediced portfolio values of Jan 01, 2020 (value of dec 31 * gross returns)
  # "gain_loss_pred": differences between predicted portoflio values and dec 31 portfolio values 
jan_1_pred_tp <- data.frame(AAPL_pred = (var_data_TP$AAPL_g_return)*dec_31_2019_tp[1,2],
                            BA_pred = (var_data_TP$BA_g_return)*dec_31_2019_tp[1,3],
                            GS_pred = (var_data_TP$GS_g_return)*dec_31_2019_tp[1,4],
                            MMM_pred = (var_data_TP$MMM_g_return)*dec_31_2019_tp[1,5],
                            GILD_pred = (var_data_TP$GILD_g_return)*dec_31_2019_tp[1,6],
                            MWHYX_pred = (var_data_TP$MWHYX_g_return)*dec_31_2019_tp[1,7],
                            IGSB_pred = (var_data_TP$IGSB_g_return)*dec_31_2019_tp[1,8],
                            VCSH_pred = (var_data_TP$VCSH_g_return)*dec_31_2019_tp[1,9],
                            IGIB_pred = (var_data_TP$IGIB_g_return)*dec_31_2019_tp[1,10],
                            VCIT_pred = (var_data_TP$VCIT_g_return)*dec_31_2019_tp[1,11])%>% 
                 mutate(portolio_value_pred = AAPL_pred*tp_shares_dec_31[1] + 
                                           BA_pred*tp_shares_dec_31[2] + 
                                           GS_pred*tp_shares_dec_31[3] + 
                                           MMM_pred*tp_shares_dec_31[4] + 
                                           GILD_pred*tp_shares_dec_31[5] +
                                           MWHYX_pred*tp_shares_dec_31[6] +
                                           IGSB_pred*tp_shares_dec_31[7] +
                                           VCSH_pred*tp_shares_dec_31[8] +
                                           IGIB_pred*tp_shares_dec_31[9] +
                                           VCIT_pred*tp_shares_dec_31[10],
                        gain_loss_pred = portolio_value_pred - dec_31_2019_tp$portfolio_value) %>%
                 na.omit()

# histogram of grain/losses
ggplot(data = jan_1_pred_tp, aes(gain_loss_pred)) + 
  geom_histogram(fill = "deepskyblue3") +
  xlab("Gain/Loss") +
  ylab("Frequency") +
  labs(title = "Histogram of Gain/Loss (Tangency Portfolio)")

# We have a total of 502 predicted gain/losses, and the 99% value at risk for one day is approximately the fifth smallest value

var_tp_1day <- sort(jan_1_pred_tp$gain_loss_pred)[5]
abs(var_tp_1day)

# calculate 15 day 99% VaR
var_tp_5day <- var_tp_1day*sqrt(5)
abs(var_tp_5day)

```


Historical Simulation Method
Calculating VaR (Minimum Variance Portfolio)
Short-Selling Not Allowed 
```{r}
# Assume that we have purchased stocks and bonds based on weights from MVR on Dec 31, 2019, here are the  stock shares
# (to see how good the weights are, we apply the weights on the day that is closest to our intended day of investment)
mvr_shares_dec_31_ns <- c((100000*mvr_noSS_weights[1])/(var_data_mvr$AAPL_AC[503]),
                         (100000*mvr_noSS_weights[2])/(var_data_mvr$BA_AC[503]),
                         (100000*mvr_noSS_weights[3])/(var_data_mvr$GS_AC[503]),
                         (100000*mvr_noSS_weights[4])/(var_data_mvr$MMM_AC[503]),
                         (100000*mvr_noSS_weights[5])/(var_data_mvr$GILD_AC[503]),
                         (100000*mvr_noSS_weights[6])/(var_data_mvr$MWHYX_AC[503]),
                         (100000*mvr_noSS_weights[7])/(var_data_mvr$IGSB_AC[503]),
                         (100000*mvr_noSS_weights[8])/(var_data_mvr$VCSH_AC[503]),
                         (100000*mvr_noSS_weights[9])/(var_data_mvr$IGIB_AC[503]),
                         (100000*mvr_noSS_weights[10])/(var_data_mvr$VCIT_AC[503]))




# calculate portfolio value of dec 31, 2019
dec_31_2019_mvr_ns <- var_data_mvr[503,] %>%
                      mutate(portfolio_value = mvr_shares_dec_31_ns[1]*AAPL_AC+
                                               mvr_shares_dec_31_ns[2]*BA_AC +
                                               mvr_shares_dec_31_ns[3]*GS_AC +
                                               mvr_shares_dec_31_ns[4]*MMM_AC +
                                               mvr_shares_dec_31_ns[5]*GILD_AC +
                                               mvr_shares_dec_31_ns[6]*MWHYX_AC +
                                               mvr_shares_dec_31_ns[7]*IGSB_AC +
                                               mvr_shares_dec_31_ns[8]*VCSH_AC +
                                               mvr_shares_dec_31_ns[9]*IGIB_AC +
                                               mvr_shares_dec_31_ns[10]*VCIT_AC) 

## use past gross returns to predict the market value of the portfolio for Jan 1, 2020 (the next day Dec 31, 2019)
  # AAPL_pred: predicted stock price
  # "portfolio_value_pred": prediced portfolio values of Jan 01, 2020 (value of dec 31 * gross returns)
  # "gain_loss_pred": differences between predicted portoflio values and dec 31 portfolio values 
jan_1_pred_mvr_ns <- data.frame(AAPL_pred = (var_data_mvr$AAPL_g_return)*dec_31_2019_mvr_ns[1,2],
                                BA_pred = (var_data_mvr$BA_g_return)*dec_31_2019_mvr_ns[1,3],
                                GS_pred = (var_data_mvr$GS_g_return)*dec_31_2019_mvr_ns[1,4],
                                MMM_pred = (var_data_mvr$MMM_g_return)*dec_31_2019_mvr_ns[1,5],
                                GILD_pred = (var_data_mvr$GILD_g_return)*dec_31_2019_mvr_ns[1,6],
                                MWHYX_pred = (var_data_mvr$MWHYX_g_return)*dec_31_2019_mvr_ns[1,7],
                                IGSB_pred = (var_data_mvr$IGSB_g_return)*dec_31_2019_mvr_ns[1,8],
                                VCSH_pred = (var_data_mvr$VCSH_g_return)*dec_31_2019_mvr_ns[1,9],
                                IGIB_pred = (var_data_mvr$IGIB_g_return)*dec_31_2019_mvr_ns[1,10],
                                VCIT_pred = (var_data_mvr$VCIT_g_return)*dec_31_2019_mvr_ns[1,11])%>%
                     mutate(portolio_value_pred = AAPL_pred*mvr_shares_dec_31_ns[1] + 
                                                  BA_pred*mvr_shares_dec_31_ns[2] + 
                                                  GS_pred*mvr_shares_dec_31_ns[3] + 
                                                  MMM_pred*mvr_shares_dec_31_ns[4] + 
                                                  GILD_pred*mvr_shares_dec_31_ns[5] +
                                                  MWHYX_pred*mvr_shares_dec_31_ns[6] +
                                                  IGSB_pred*mvr_shares_dec_31_ns[7] +
                                                  VCSH_pred*mvr_shares_dec_31_ns[8] +
                                                  IGIB_pred*mvr_shares_dec_31_ns[9] +
                                                  VCIT_pred*mvr_shares_dec_31_ns[10],
                     gain_loss_pred = portolio_value_pred - dec_31_2019_mvr_ns$portfolio_value) %>%
                     na.omit()


# histogram of grain/losses
ggplot(data = jan_1_pred_mvr_ns, aes(gain_loss_pred)) + 
  geom_histogram(fill = "deepskyblue3") +
  xlab("Gain/Loss") +
  ylab("Frequency") +
  labs(title = "Histogram of Gain/Loss (Minimum Variance Portfolio: No Short-Selling)")

# We have a total of 502 predicted gain/losses, and the 99% value at risk for one day is approximately the fifth smallest value

var_mvr_1day_ns <- sort(jan_1_pred_mvr_ns$gain_loss_pred)[5]
abs(var_mvr_1day_ns)

# calculate 15 day 99% VaR
var_mvr_5day_ns <- var_mvr_1day_ns*sqrt(5)
abs(var_mvr_5day_ns)


```


Historical Simulation Method
Calculating VaR (Tangnecy Portfolio)
Short-Selling Not Allowed 
```{r}
# Assume that we have purchased stocks based on weights from TP on Dec 31, 2019, here are the stock shares
  # (notice that some shares are negative)
tp_shares_dec_31_ns <- c((100000*t_noSS_weights[1])/(var_data_TP$AAPL_AC[503]),
                        (100000*t_noSS_weights[2])/(var_data_TP$BA_AC[503]),
                        (100000*t_noSS_weights[3])/(var_data_TP$GS_AC[503]),
                        (100000*t_noSS_weights[4])/(var_data_TP$MMM_AC[503]),
                        (100000*t_noSS_weights[5])/(var_data_TP$GILD_AC[503]),
                        (100000*t_noSS_weights[6])/(var_data_TP$MWHYX_AC[503]),
                        (100000*t_noSS_weights[7])/(var_data_TP$IGSB_AC[503]),
                        (100000*t_noSS_weights[8])/(var_data_TP$VCSH_AC[503]),
                        (100000*t_noSS_weights[9])/(var_data_TP$IGIB_AC[503]),
                        (100000*t_noSS_weights[10])/(var_data_TP$VCIT_AC[503]))


# calculate portfolio value of dec 31, 2019

dec_31_2019_tp_ns <- var_data_TP[503,] %>%
                     mutate(portfolio_value = 
                                        tp_shares_dec_31_ns[1]*AAPL_AC+
                                        tp_shares_dec_31_ns[2]*BA_AC +
                                        tp_shares_dec_31_ns[3]*GS_AC +
                                        tp_shares_dec_31_ns[4]*MMM_AC +
                                        tp_shares_dec_31_ns[5]*GILD_AC +
                                        tp_shares_dec_31_ns[6]*MWHYX_AC +
                                        tp_shares_dec_31_ns[7]*IGSB_AC +
                                        tp_shares_dec_31_ns[8]*VCSH_AC +
                                        tp_shares_dec_31_ns[9]*IGIB_AC +
                                        tp_shares_dec_31_ns[10]*VCIT_AC) 

## use past gross returns to predict the market value of the portfolio for Jan 1, 2020 (the next day Dec 31, 2019)
  # AAPL_pred: predicted stock price 
  # "portfolio_value_pred": prediced portfolio values of Jan 01, 2020 (value of dec 31 * gross returns)
  # "gain_loss_pred": differences between predicted portoflio values and dec 31 portfolio values 
jan_1_pred_tp_ns <- data.frame(AAPL_pred = (var_data_TP$AAPL_g_return)*dec_31_2019_tp_ns[1,2],
                               BA_pred = (var_data_TP$BA_g_return)*dec_31_2019_tp_ns[1,3],
                               GS_pred = (var_data_TP$GS_g_return)*dec_31_2019_tp_ns[1,4],
                               MMM_pred = (var_data_TP$MMM_g_return)*dec_31_2019_tp_ns[1,5],
                               GILD_pred = (var_data_TP$GILD_g_return)*dec_31_2019_tp_ns[1,6],
                               MWHYX_pred = (var_data_TP$MWHYX_g_return)*dec_31_2019_tp_ns[1,7],
                               IGSB_pred = (var_data_TP$IGSB_g_return)*dec_31_2019_tp_ns[1,8],
                               VCSH_pred = (var_data_TP$VCSH_g_return)*dec_31_2019_tp_ns[1,9],
                               IGIB_pred = (var_data_TP$IGIB_g_return)*dec_31_2019_tp_ns[1,10],
                               VCIT_pred = (var_data_TP$VCIT_g_return)*dec_31_2019_tp_ns[1,11])%>% 
                    mutate(portolio_value_pred = AAPL_pred*tp_shares_dec_31_ns[1] + 
                                                 BA_pred*tp_shares_dec_31_ns[2] + 
                                                 GS_pred*tp_shares_dec_31_ns[3] + 
                                                 MMM_pred*tp_shares_dec_31_ns[4] + 
                                                 GILD_pred*tp_shares_dec_31_ns[5] +
                                                 MWHYX_pred*tp_shares_dec_31_ns[6] +
                                                 IGSB_pred*tp_shares_dec_31_ns[7] +
                                                 VCSH_pred*tp_shares_dec_31_ns[8] +
                                                 IGIB_pred*tp_shares_dec_31_ns[9] +
                                                 VCIT_pred*tp_shares_dec_31_ns[10],
                        gain_loss_pred = portolio_value_pred - dec_31_2019_tp_ns$portfolio_value) %>%
                    na.omit()

# histogram of grain/losses
ggplot(data = jan_1_pred_tp_ns, aes(gain_loss_pred)) + 
  geom_histogram(fill = "deepskyblue3") +
  xlab("Gain/Loss") +
  ylab("Frequency") +
  labs(title = "Histogram of Gain/Loss (Tangency Portfolio): No Short-Selling")

# We have a total of 502 predicted gain/losses, and the 99% value at risk for one day is approximately the fifth smallest value

var_tp_1day_ns <- sort(jan_1_pred_tp_ns$gain_loss_pred)[5]
abs(var_tp_1day_ns)

# calculate 15 day 99% VaR
var_tp_5day_ns <- var_tp_1day_ns*sqrt(5)
abs(var_tp_5day_ns)
```

